---
title: "introduction_to_empirical_bayes"
author: "Dylan"
date: "13/04/2022"
output: pdf_document
---

```{r setup}
library(Lahman) 
library(dplyr)
library(tidyr)
library(ebbr)
library(tidyverse)
```


Grab career batting average of non-pitchers 
(allow players that have pitched <= 3 games, like Ty Cobb)
```{r pitchers}
pitchers <- Pitching %>% 
  group_by(playerID) %>% 
  summarize(gamesPitched = sum(G)) %>% 
  filter(gamesPitched > 3)

player_names <- Master %>%
  tbl_df () %>% 
  select(playerID, nameFirst, nameLast, bats) %>% 
  unite(name, nameFirst, nameLast, sep = " ")

career_full <- Batting %>% 
  filter(AB > 0) %>% 
  anti_join(pitchers, by = "playerID") %>% 
  group_by(playerID) %>% 
  summarize(H = sum(H), AB = sum(AB), year = mean(yearID)) %>% 
  inner_join(player_names, by = "playerID") %>% 
  filter(!is.na(bats)) 

career <- career_full %>% 
  select(-bats, -year)
```


Empirical Bayes estimation 
```{r ebe}
prior <- career %>% 
  filter(AB >= 500) %>% 
  ebb_fit_prior(H, AB)

eb_career <- career %>% 
  add_ebb_estimate(H, AB, prior_subset = AB >= 500)

eb_career %>% 
  ggplot(aes(.raw, .fitted, color = AB)) + 
  geom_point() + 
  geom_abline(color = "red") + 
  scale_color_continuous(trans = "log", breaks = c(1,10,100,1000)) + 
  geom_hline(yintercept = tidy(prior)$mean, color = "red", lty = 2) + 
  labs(x = "Raw batting average",
       y = "Shruken batting average")

yankee_1998 <- c("brosisc01", "jeterde01", "knoblch01", "martiti02", "posadjo01", "strawda01", "willibe02")
eb_career %>% 
  filter(playerID %in% yankee_1998) %>% 
  mutate(name = reorder(name, .fitted)) %>% 
  ggplot(aes(.fitted, name)) + 
  geom_point() + 
  geom_errorbarh(aes(xmin = .low, xmax = .high)) + 
  labs(x = "Estimated batting average (w/ 95% confidence interval",
       y = "Player")
```


Hierarchical modeling 
```{r hm}
career %>% 
  filter(AB >= 10) %>% 
  ggplot(aes(AB, H / AB)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  scale_x_log10() 

eb_career_ab <- career %>% 
  add_ebb_estimate(H, AB, method = "gamlss",
                   mu_predictors = ~ log10(AB))

eb_career_ab

eb_career_ab %>%
  filter(AB > 10) %>%
  rename(Raw = .raw, Shrunken = .fitted) %>% gather(type, estimate, Raw, Shrunken) %>% ggplot(aes(AB, estimate)) +
  geom_point() +
  facet_wrap(~ type) +
  scale_x_log10()
```










